---
title: Writing Functions
author: John Compton
date: '2020-03-09'
slug: writing-functions
categories: []
tags:
  - r
  - functions
  - data visualization
draft: false
---

In Stat 385 one of the main things we've been working on is writing various and more complicated functions. 

For example we wrote the following vectorized roulette function to simulate the rewards for a French Roulette.

A French roulette has:
  37 colored and numbered pockets on the wheel.
  0 is green.
  In number ranges from 1 to 10 and 19 to 28:
  Odd numbers are red.
  Even numbers are black.
  In ranges from 11 to 18 and 29 to 36:
  Odd numbers are black.
  Even numbers are red.

```{r}
roulette_vec = function(many_bets)
{
  win_lose_random <- sample(x = c(TRUE, FALSE), size = length(many_bets),
                            replace = TRUE)
  many_bets[many_bets == "high"|| many_bets == "low"] = 10
  many_bets[many_bets == "red"|| many_bets == "black"] = 20
  many_bets[many_bets == "odd"|| many_bets == "even"] = 15
  many_bets[many_bets == "first"|| many_bets == "second" || many_bets == "third"] = 50
  ifelse(many_bets[win_lose_random], many_bets, 0)
}
```


We have also been working on data visualizations:

```{r}
hist(faithful$eruptions, breaks = 18, border = "dodgerblue", probability = TRUE, xlim = c(1,6), 
     ylim = c(0, 0.85), xlab = "Eruption Time (mins)", main = "Histogram of Old Faithful Geyser Eruption Time")
box()
grid()
lines(density(faithful$eruptions), lwd = 2, col = "Orange")
```
